# Does body composition affect wearable signal reliability?
## PAMAP2 Physical Activity Monitoring (UCI ML Repository, id=231)

```{r}
library(lme4)
library(sandwich)
library(lmtest)
library(MASS)

pamap2 <- read.csv("~/pamap2/pamap2_quality_clean.csv")
str(pamap2)
nrow(pamap2)
```

### Exploratory: subject-level, per-location patterns
```{r}
subj.loc.avg <- aggregate(variance ~ subject_id + location, data = pamap2, FUN = mean)
bmi.lookup <- unique(pamap2[, c("subject_id", "bmi")])
subj.loc.avg <- merge(subj.loc.avg, bmi.lookup, by = "subject_id")
 
locations <- unique(subj.loc.avg$location)
 
cat("\n Correlation: BMI vs. variance, by location \n")
for (loc in locations) {
  sub <- subj.loc.avg[subj.loc.avg$location == loc, ]
  cor.test.result <- cor.test(sub$bmi, sub$variance)
  cat(loc, ": r=", round(cor.test.result$estimate, 3),
      ", p=", round(cor.test.result$p.value, 3),
      " (n=", nrow(sub), ")\n", sep = "")
}
 
cat("\n ANOVA: variance ~ location \n")
anova.fit <- aov(variance ~ location, data = pamap2)
print(summary(anova.fit))
 
par(mfrow = c(1, 3))
for (loc in locations) {
  sub <- subj.loc.avg[subj.loc.avg$location == loc, ]
  sub <- sub[is.finite(sub$bmi) & is.finite(sub$variance), ]
  if (nrow(sub) < 2) {
    cat("Skipping plot for", loc, "-- not enough valid data points\n")
    next
  }
  plot(sub$bmi, sub$variance, pch = 16, main = loc,
       xlab = "BMI", ylab = "Average Signal Variance")
  loc.fit <- lm(variance ~ bmi, data = sub)
  abline(loc.fit, col = "deeppink")
}
```

### Multivariate Model
Pooling all 273 rows treats each subject's ~30 readings as independent, which understates uncertainty on bmi (bmi only truly varies across 9 subjects). 
```{r}
ols.fit <- lm(variance ~ bmi + location + activity, data = pamap2)
cat("\n Naive OLS (baseline, not final): bmi coefficient \n")
print(coeftest(ols.fit, vcov = vcovHC(ols.fit, type = "HC1"))["bmi", ])
```

### Mixed-effects model
```{r}
mixed.fit <- lmer(variance ~ bmi + location + activity + (1 | subject_id), data = pamap2)
cat("\n Mixed-effects model (subject as random effect) \n")
coefs <- summary(mixed.fit)$coefficients
p.values <- 2 * (1 - pnorm(abs(coefs[, "t value"])))
mixed.results <- data.frame(
  Estimate = round(coefs[, "Estimate"], 3),
  t.value = round(coefs[, "t value"], 3),
  p.value = round(p.values, 3)
)
print(mixed.results)

# BMI's main effect assumes one constant effect across all location but the real question is whether BMI's effect differs by location.
mixed.fit.interaction <- lmer(variance ~ bmi * location + activity + (1 | subject_id), data = pamap2)
cat("\n Mixed-effects model WITH bmi x location interaction \n")
print(summary(mixed.fit.interaction)$coefficients)
```

### Robustness
```{r}

# 4a. Robust regression (downweights outliers automatically)
robust.fit <- rlm(variance ~ bmi + location + activity, data = pamap2)
cat("\n Robust regression: bmi coefficient \n")
print(summary(robust.fit)$coefficients["bmi", ])
 
# 4b. Bootstrapped CI on the BMI-variance correlation, by location
set.seed(42)
n.boot <- 2000
boot.results <- data.frame(location = character(), r = numeric(), ci.lower = numeric(), ci.upper = numeric())
for (loc in locations) {
  sub <- subj.loc.avg[subj.loc.avg$location == loc, ]
  sub <- sub[is.finite(sub$bmi) & is.finite(sub$variance), ]
  boot.cors <- replicate(n.boot, {
    resample.idx <- sample(1:nrow(sub), size = nrow(sub), replace = TRUE)
    resampled <- sub[resample.idx, ]
    if (length(unique(resampled$bmi)) < 2) return(NA)
    cor(resampled$bmi, resampled$variance)
  })
  boot.cors <- boot.cors[!is.na(boot.cors)]
  boot.results <- rbind(boot.results, data.frame(
    location = loc,
    r = round(cor(sub$bmi, sub$variance), 3),
    ci.lower = round(quantile(boot.cors, 0.025), 3),
    ci.upper = round(quantile(boot.cors, 0.975), 3)
  ))
}
rownames(boot.results) <- NULL
boot.results

write.csv(boot.results, "~/pamap2/bootstrap_ci.csv", row.names = FALSE)
 
# 4c. Leave-one-subject-out sensitivity 
cat("\n Leave-one-out: ANOVA p-value (variance ~ location) \n")
subjects <- unique(pamap2$subject_id)
loo.results <- data.frame(excluded_subject = subjects, p.value = NA)
for (i in seq_along(subjects)) {
  pamap2.subset <- pamap2[pamap2$subject_id != subjects[i], ]
  fit <- aov(variance ~ location, data = pamap2.subset)
  loo.results$p.value[i] <- summary(fit)[[1]][["Pr(>F)"]][1]
}
print(loo.results)
cat("Range of p-values across leave-one-out refits:",
    round(min(loo.results$p.value), 3), "to", round(max(loo.results$p.value), 3), "\n")
```

### Summary: BMI effect across specifications
```{r}
summary.table <- data.frame(
  Model = c(
    "Correlation (hand only)",
    "Naive OLS (baseline)",
    "Robust regression (rlm)",
    "Mixed-effects (main effect)"
  ),
  Estimate = c(
    cor(subj.loc.avg$bmi[subj.loc.avg$location == "hand"],
        subj.loc.avg$variance[subj.loc.avg$location == "hand"]),
    coef(ols.fit)["bmi"],
    coef(robust.fit)["bmi"],
    fixef(mixed.fit)["bmi"]
  )
)
print(summary.table)
```
```{r}
write.csv(boot.results, "~/pamap2/bootstrap_ci.csv", row.names = FALSE)
```

