# Does body composition affect wearable signal reliability?
## PAMAP2 Physical Activity Monitoring (UCI ML Repository, id=231)

```{r}
library(sandwich)
library(lmtest)
library(MASS)

pamap2 <- read.csv("~/pamap2/pamap2_quality_clean.csv")
str(pamap2)
nrow(pamap2)
```

### Exploratory: subject-level, per-location patterns
```{r}

subj.loc.avg <- aggregate(variance ~ subject_id + location, 
                          data = pamap2, FUN = mean)
bmi.lookup <- unique(pamap2[, c("subject_id", "bmi")])
subj.loc.avg <- merge(subj.loc.avg, bmi.lookup, by = "subject_id")
 
locations <- unique(subj.loc.avg$location)
 
# correlation: BMI vs. variance, by location
cor.results <- data.frame(location = character(), r = numeric(), 
                          p.value = numeric(), n = integer())
for (loc in locations) {
  sub <- subj.loc.avg[subj.loc.avg$location == loc, ]
  cor.test.result <- cor.test(sub$bmi, sub$variance)
  cor.results <- rbind(cor.results, data.frame(
    location = loc,
    r = round(cor.test.result$estimate, 3),
    p.value = round(cor.test.result$p.value, 3),
    n = nrow(sub)
  ))
}
cor.results
 
# does variance differ by location overall?
anova.fit <- aov(variance ~ location, data = pamap2)
summary(anova.fit)
 
par(mfrow = c(1, 3))
for (loc in locations) {
  sub <- subj.loc.avg[subj.loc.avg$location == loc, ]
  sub <- sub[is.finite(sub$bmi) & is.finite(sub$variance), ]
  if (nrow(sub) < 2) next
  plot(sub$bmi, sub$variance, pch = 16, main = loc,
       xlab = "BMI", ylab = "Average Signal Variance")
  loc.fit <- lm(variance ~ bmi, data = sub)
  abline(loc.fit, col = "deeppink")
}
```

### Multivariate OLS with robust standard errors
Pooling all 273 rows treats each subject's ~30 readings as independent, which is why the per-location correlations above (computed on subject-level averages) are the more trustworthy test of BMI's effect (this pooled model is included as a baseline for comparison).
```{r}
ols.fit <- lm(variance ~ bmi + location + activity, data = pamap2)
coeftest(ols.fit, vcov = vcovHC(ols.fit, type = "HC1"))["bmi", ]
```


### Robustness check
```{r}
# robust regression
robust.fit <- rlm(variance ~ bmi + location + activity, data = pamap2)
summary(robust.fit)$coefficients["bmi", ]

# bootstrapped 95% CI on the BMI-variance correlation, by location. Because n=9 is small, the usual formula-based CI for a correlation assumes a larger sample than we have. Bootstrapping instead resamples our own 9 subjects (with replacement) many times to see how much the correlation would plausibly bounce around with a different sample of 9.

set.seed(42)
n.boot <- 2000
boot.results <- data.frame(location = character(), r = numeric(), ci.lower = numeric(), ci.upper = numeric())
for (loc in locations) {
  sub <- subj.loc.avg[subj.loc.avg$location == loc, ]
  sub <- sub[is.finite(sub$bmi) & is.finite(sub$variance), ]
  boot.cors <- replicate(n.boot, {
    resample.idx <- sample(1:nrow(sub), size = nrow(sub), replace = TRUE)
    resampled <- sub[resample.idx, ]
    if (length(unique(resampled$bmi)) < 2) return(NA)
    cor(resampled$bmi, resampled$variance)
  })
  boot.cors <- boot.cors[!is.na(boot.cors)]
  boot.results <- rbind(boot.results, data.frame(
    location = loc,
    r = round(cor(sub$bmi, sub$variance), 3),
    ci.lower = round(quantile(boot.cors, 0.025), 3),
    ci.upper = round(quantile(boot.cors, 0.975), 3)
  ))
}
rownames(boot.results) <- NULL
boot.results
 
write.csv(boot.results, "bootstrap_ci.csv", row.names = FALSE)
 
# leave-one-subject-out: does the location effect (ANOVA) depend on any single subject?
subjects <- unique(pamap2$subject_id)
loo.results <- data.frame(excluded_subject = subjects, p.value = NA)
for (i in seq_along(subjects)) {
  pamap2.subset <- pamap2[pamap2$subject_id != subjects[i], ]
  fit <- aov(variance ~ location, data = pamap2.subset)
  loo.results$p.value[i] <- summary(fit)[[1]][["Pr(>F)"]][1]
}
loo.results
range(loo.results$p.value)
```

## Summary: BMI effect across specifications
```{r}

summary.table <- data.frame(
  Model = c(
    "Correlation (hand only)",
    "Naive pooled OLS",
    "Robust regression (rlm)"
  ),
  Estimate = c(
    cor(subj.loc.avg$bmi[subj.loc.avg$location == "hand"],
        subj.loc.avg$variance[subj.loc.avg$location == "hand"]),
    coef(ols.fit)["bmi"],
    coef(robust.fit)["bmi"]
  )
)
summary.table
```

